---
title: "eBird India monthly report - November 2016"
output: html_document
---

<style>
  .col2 {
    columns: 2 250px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 250px; /* chrome, safari */
    -moz-columns: 2 250px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>

```{r data, ignore=TRUE, include=FALSE}
library(dplyr)
library(ggplot2)
library(rgdal)
library(raster)
library(RColorBrewer)
library(knitr)
library(pander)


load("C:/Users/Abinand Reddy/Desktop/Thesis/NCF/eBird/All-eBird-India-data-2016-12-06.RData")

all$observer <- gsub('[[:alpha:]]', '', all$OBSERVER.ID)

states <- read.csv("C:/Users/Abinand Reddy/Desktop/Thesis/NCF/eBird/state_codes.csv")

ind <- readOGR('C:/Users/Abinand Reddy/Desktop/Thesis/NCF/eBird/District shape file eBird code', layer = 'Distrits with eBird code')

thismonth <- '2016-11'
lastyearmonth <- '2015-11'
lastmonth <- '2016-10'
nextmonth <- '2016-12'
threemonths <- '2016-09'

```

##Number of lists

```{r, echo=FALSE, warning=F}
number_lists <- all %>% filter(year.month > '2013-12' & year.month <= thismonth) %>% group_by(year.month) %>% summarise(n_lists = n_distinct(SAMPLING.EVENT.IDENTIFIER))


number_lists$year <- do.call('rbind',strsplit(number_lists$year.month, '-'))[,1]

number_lists$month <- do.call('rbind',strsplit(number_lists$year.month, '-'))[,2]

number_lists1 <- with(number_lists, expand.grid(month = levels(factor(number_lists$month)), year = levels(factor(number_lists$year))))

number_lists2 <- merge(number_lists,number_lists1, all.y = T)


l <- ggplot(number_lists2, aes(x = month, y = n_lists, color = year, group = year, shape = year))+
  geom_line() +
  geom_point() +
  ylab('Number of lists')+
  theme_minimal()
l

```

## Users per State - November 2016

```{r users, echo=FALSE, fig.height = 8, results= 'asis'}

state_wise_users <- all %>% group_by(SUBNATIONAL1_CODE,upload.month) %>% summarise(npeople = n_distinct(observer)) 

state_wise_two_months_users <- state_wise_users %>% filter(upload.month == lastyearmonth| upload.month == thismonth)

state_wise_two_months_users$state <- states$state[match(state_wise_two_months_users$SUBNATIONAL1_CODE, states$SUBNATIONAL1_CODE)]

state_wise_two_months_users1 <- with(state_wise_two_months_users,expand.grid(state = levels(factor(states$state)), upload.month = levels(factor(upload.month))))

state_wise_two_months_users2 <- merge(state_wise_two_months_users,state_wise_two_months_users1,all.y = T)

state_wise_two_months_users3 <- state_wise_two_months_users2 %>% arrange(upload.month, npeople)


state_wise_two_months_users3$npeople[is.na(state_wise_two_months_users3$npeople)== T] <- 0

p <- ggplot(state_wise_two_months_users3, aes(x = reorder(state, npeople), y= npeople, fill = upload.month))+
  geom_bar(stat = 'identity', position = position_dodge(), colour = 'black', width = 0.75) +
  scale_fill_manual(values = c('white','light blue'), labels = c('2015','2016'))+
  guides(fill = guide_legend(title = ''))+
  scale_y_continuous(sec.axis = dup_axis())+
  ylab('Users') +
  xlab('')+
  coord_flip()+
  theme_minimal()

p


```



##Lists per State - November 2016


```{r lists, warning=FALSE, echo = FALSE,  fig.height = 8}

state_wise_lists <- all %>% group_by(SUBNATIONAL1_CODE,year.month, ALL.SPECIES.REPORTED == 1) %>% summarise(clists = n_distinct(SAMPLING.EVENT.IDENTIFIER)) %>% group_by(SUBNATIONAL1_CODE,year.month) %>% mutate(nlists = sum(clists)) %>% filter(`ALL.SPECIES.REPORTED == 1` == T)

state_wise_two_months_lists <- state_wise_lists %>% filter(year.month == lastyearmonth| year.month == thismonth)

state_wise_two_months_lists$state <- states$state[match(state_wise_two_months_lists$SUBNATIONAL1_CODE, states$SUBNATIONAL1_CODE)]

state_wise_two_months_lists1 <- with(state_wise_two_months_lists,expand.grid(state = levels(factor(states$state)), year.month = levels(factor(year.month))))

state_wise_two_months_lists2 <- merge(state_wise_two_months_lists,state_wise_two_months_lists1,all.y = T)

state_wise_two_months_lists3 <- state_wise_two_months_lists2 %>% arrange(year.month, nlists)

state_wise_two_months_lists3$nlists[is.na(state_wise_two_months_lists3$nlists)== T] <- 0

state_wise_two_months_lists3$clists[state_wise_two_months_lists3$year.month == lastyearmonth] <- NA


p <- ggplot(state_wise_two_months_lists3, aes(x = reorder(state, nlists), y = nlists, fill = year.month))+
  geom_bar(stat = 'identity', position = position_dodge(), colour = 'black', width = 0.75) +
  scale_fill_manual(values = c('white','coral'), labels = c('2015','2016'))+
  geom_text(aes(x = reorder(state, nlists), y = nlists + 300, label = clists))+
  guides(fill = guide_legend(title = ''))+
  scale_y_continuous(sec.axis = dup_axis())+
  ylab('lists') +
  xlab('')+
  coord_flip()+
  theme_minimal()
p

```

The graph shows the number of lists uploaded by users per state. Shared lists are counted as multiple lists uploaded by multiple users and not as a single list. Numbers next to bars represent the number of the lists, that are complete.

##District Covereage per State - November 2016

```{r, fig.height=8, warning=FALSE, echo=F}
dist <- ind@data

dist <- dist %>% group_by(ST_NM) %>% summarise(n_dist = n_distinct(district10))

numd <- all %>% filter(year.month %in% c(thismonth,lastyearmonth)) %>% group_by(year.month,SUBNATIONAL1_CODE) %>% filter(!is.na(SUBNATIONAL2_CODE)) %>% summarise(dis = n_distinct(SUBNATIONAL2_CODE))



numd$state <- states$state[match(numd$SUBNATIONAL1_CODE, states$SUBNATIONAL1_CODE)]

numd <- left_join(numd,dist, by = c("state"="ST_NM"))


numd1 <- with(numd,expand.grid(state = levels(factor(states$state)), year.month = levels(factor(year.month))))

colnames(numd)[4] <- 'state'

numd2 <- merge(numd,numd1,all.y = T)

numd2 <- numd2 %>% mutate(cov = round(dis/n_dist,2))

numd2$cov[is.na(numd2$cov)== T] <- 0

numd2 <- numd2[!numd2$state == 'India',]


p <- ggplot(numd2, aes(x = reorder(state, cov), y = cov, fill = year.month))+
  geom_bar(stat = 'identity', position = position_dodge(), colour = 'black', width = 0.75) +
  scale_fill_manual(values = c('white','darkolivegreen3'), labels = c('2015','2016'))+
  guides(fill = guide_legend(title = ''))+
  scale_y_continuous(breaks = c(0,.25,.50,.75,1),labels = scales::percent,sec.axis = dup_axis())+
  ylab('Percentage of Districts Covered') +
  xlab('')+
  coord_flip()+
  theme_minimal()
p


```



##User Activity - November 2016

```{r User activity, echo=FALSE, warning=FALSE}
all1 <- all %>% filter(upload.month < nextmonth) %>% group_by(observer) %>% mutate(f_upload = min(upload.month,na.rm = T))


us_act <- rbind(c('New Users November 2016', length(unique(all1$observer[all1$f_upload == thismonth]))),
      c('All Users November 2016', length(unique(all1$observer[all1$upload.month == thismonth]))),
      c('All Users since September 2016 (Past 3 months)', length(unique(all1$observer[all1$upload.month >= threemonths]))),
      c('All Users', length(unique(all1$observer))))


colnames(us_act) <- c('Activity','Users')

kable(us_act)

```





**Data downloaded**: 2016-12-06